---
title: "County Modelling"
author: "Iben M. Ricket"
date: "9/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#County Modelling Part 5: Post model evaluation

Best performing models were identified in a systematic capacity using MSE. The best performing models, for each outcome-metric, were from iteration 4, used random forest as the feature selection technique, and utilized a non-parametric machine learning algorithm. The best performing models for each outcome metric were as follows: 

(1) ER: Random Forest (MSE: 0.003, R2: 0.279)
(2) Inpatient days: Gradient Boosting Machines (MSE: 0.009, R2: 0.267)
(3) Hospital expenditures (MSE: 0.003, R2: 0.447)

Best performing models were used in several post model evaluation exercises. 

Libraries needed to support analysis 
```{r message=FALSE, warning=FALSE}
library(healthcareai)
library(dplyr)
library(glmnet)
library(pROC)
library(caret)
library(ROCR)
library(generalhoslem)
library(ResourceSelection)
library(PredictABEL)
library(ROCR)
library(SDMTools)
library(randomForest)
library(arsenal)
library(sqldf)
library(tidyverse)
library(DAAG)
library(sl3)
library(gridExtra)
```


